Distributed Topology Discovery in Self-Assembled Nano Network-On-Chip

### 2014-09-09 PAPER/implementation.tex
Fixed bio for each author and some other typos

A possible hardware implementation of \disr{} algorithm will be
described in this section. Figure~\ref{fig:node_structure} depicts the general
structure of a node in the particular scenario of DNA nano Network-On-Chip. In particular such a node
is composed by the following fundamental elements:

@@

To give a basic estimate of the overhead needed we focus on \disr{}-
specific components namely the control logic and configuration
registers needed to implement \disr{}. Figure~\ref{fig:implementation}
shows an architecture sketch, which mainly consist in the
following building elements:

@@

	\item \textbf{Packager:}
	Essentially it is a set of registers and multiplexers. Starting from
	an incoming packet stored in the input buffer, it updates packet's
	content before retransmission to a destination
	node.  In some cases, this circuitry builds the packet from
	scratch e.g. when a node is bootstrap and for
	the first time it should inject a \emph{starting\_segment\_request}. 
	Figure~\ref{fig:packager} depicts the architecture of this
	block. When a packet is created for the first time, the multiplexers 
	driven by the controller, will select information incoming from an internal 
    node such as the \emph{source\_ID}. In this situation the 
	TTL field should be reset to its initial value and the \emph{segment\_ID}
	is composed by the node ID and by the output port identification 
	number.

@@

All the blocks described above, reacts with the rising edge of
the \emph{clock signal}. An asynchronous system \emph{reset} signal is
also present to restore all registers to their default values. When
this signal is set, DBS state machines for each network
node are set as \emph{Free}.  
Figure~\ref{fig:pipeline} is shows the pipeline structure of the
\disr{} architecture. Packets incoming from neighbor nodes (or
generated locally) are sent to the DBS state machine in order to
ensure the status commutation. While DBS updates its state, in the
next clock cycle the control circuitry will update LED registers status and
will drive the \emph{Switch Matrix} in order to send eventually the packets to
the right output buffer.
%--------------------------------------------------------------------
\subsection{Synthesis results: Timing, Area and Power consumption considerations}

As discussed above, one of the main design challenges of DNA Self-
Assembled systems is the limited number of resources available to implement both
computations and routing decisions for each node. Assuming a budget of
$10^4$ CNFETs for each network's node~\cite{liu_jetcs}  we estimated
the required resources to implement the entire \disr{} block. The RTL
description of the circuitry described in the last section (with an
hardware description language)  of the \disr{} circuitry has been written
and synthesized at gate-level using \emph{Synopsys Design Compiler} with 
a generic technology library (GTECH).
Considering the specific layout of each single logic elements  (NAND,
full-adder, latch etc.), it has been possible to get a rough estimate
of the number of transistors necessary to implement \disr{} logic.
Figure~\ref{fig:synthesys} shows the results of synthesis in terms of
number of devices (CNFETs) versus the number of network nodes while
the network scales up from $10\times10$ to $100\times100$ nodes.

@@

In particular, Figure~\ref{fig:synthesys} shows that the proposed 
implementation occupies about 17,5\% of our node budget. 
The main contribution is due to the controller circuitry followed by LED registers. 
Further, more than the
absolute number of devices itself, it is interesting to observe that
the circuitry complexity which implements \disr{} algorithm increases
with a slowly growing trend. An intuitive explanation of this is the
relatively simple logic of \disr{}, whose behavior is almost coded on
scalable storage structures. For example, the number of registers
implementing the \emph{link\_visited[]} and \emph{link\_tvisited[]}
table follow the logarithmic function $N_{reg}=N_{port} \cdot
log_2(N)$ where \emph{Nport} is the number of the router’s ports and N is
the number of network’s nodes.

In order to have an idea of the maximum working clock frequency for
the implemented circuitry, timing results can be obtained from a 
gate-level netlist. Since a complete technology library useful for
commercial synthesis tool (e.g. Design Compiler) is not available for
this particular technology, we have
synthesized the RTL description of \disr{}, with a standard 32~nm CMOS
library, obtaining a delay of about $\mathrm{\tau_{\disr{}}=10} \ FO4$
(fan-out of four). Considering the results obtained
in~\cite{deng_isscc07}, which reports the ratio in terms of FO4
between a standard 32~nm CMOS technology and a carbon nanotube ones, 
rough delay estimation can be obtained. In particular in such work 
has been reported a ratio of:  
\begin{equation}
\mathrm{\frac{FO4_{CMOS}}{FO4_{CNFET}}=2}  
\end{equation}
Considering that F04 CMOS is about 30~ps, and the results of last
equation, a FO4 for the CNFET case is equal to about 15~ps, the working 
frequency can be computed as:
\begin{equation}
\mathrm{f_{clk}=\frac{1}{T_{max}}=\frac{1}{\tau_{\disr{}} \cdot FO4_{CNFET}}}=
\frac{1}{10\cdot 15 \cdot 10^{-12}}=6,6 GHz
\end{equation}
where $\mathrm{\tau_{\disr{}}}$ is expressed in terms of FO4.

Regarding the power consumption of whole \disr{} devices, 
we avoided an accurate power analysis with the switch activity
annotation due to the fact that \disr{} circuitry will be active 
only once during system startup. Indeed, after the setup phase, 
when all segments are mapped, the \disr{} block will stop its activity 
passing all the information related to segments to the routing
algorithm. For this reason the dynamic power falls to zero and
the only consumption could be due to the leakage
consumption. Finally, with the reported results, a coparison between 
RPF configuration circuits and the proposed circuits could be performed.
In particular, the state of the art implementation for the RPF algorithm has been shown in~\cite{patwardhan_nanoarch06}. As reported in this work,  the implemented hardware requires an overhead of about 1692 CNFET for a $100 \times 100$ network. Furthemore, regarding the power consumption, the RPF approach has a setup phase beyond which, the consumptions tends to zero. Starting from these results, an appreciable discrepancy in terms of transistors cost and power consumption can not be noted.

>>>>

A possible hardware implementation of \disr{} algorithm will be
described in this section. Figure~\ref{fig:node_structure} depicts the general
structure of a node in the particular scenario of DNA nano Network-On-Chip. In particular, such a node
is composed of the following fundamental elements:

@@

To give a basic estimate of the overhead needed, we focus on \disr{}-
specific components namely the control logic and configuration
registers needed to implement \disr{}. Figure~\ref{fig:implementation}
shows an architecture sketch, which mainly consists in the
following building elements:

@@

	\item \textbf{Packager:}
	Essentially it is a set of registers and multiplexers. Starting from
	an incoming packet stored in the input buffer, it updates packet's
	content before retransmission to a destination
	node.  In some cases, this circuitry builds the packet from
	scratch e.g. when a node is bootstrap and for
	the first time it should inject a \emph{STARTING\_SEGMENT\_REQUEST}. 
	Figure~\ref{fig:packager} depicts the architecture of this
	block. When a packet is created for the first time, the multiplexers 
	driven by the controller, will select information incoming from an internal 
    node such as the \emph{src\_id}. In this situation the 
	TTL field should be reset to its initial value and the \emph{segID}
	is composed of the node ID and by the output port identification 
	number.

@@

All the blocks described above, react with the rising edge of
the \emph{clock signal}. An asynchronous system \emph{reset} signal is
also present to restore all registers to their default values. When
this signal is set, DBS state machines for each network
node are set as \emph{Free}.  
Figure~\ref{fig:pipeline} shows the pipeline structure of the
\disr{} architecture. Packets incoming from neighbor nodes (or
generated locally) are sent to the DBS state machine in order to
ensure the status commutation. While DBS updates its state, in the
next clock cycle the control circuitry will update LED registers status and
will drive the \emph{Switch Matrix} in order to send the packets to
the right output buffer.
%--------------------------------------------------------------------
\subsection{Synthesis results: Timing, Area and Power consumption considerations}

As discussed above, one of the main design challenges of DNA Self-
Assembled systems is the limited number of resources available to implement both
computations and routing decisions for each node. Assuming a budget of
$10^4$ CNFETs for each network's node~\cite{liu_jetcs}  we estimated
the resources required to implement the entire \disr{} block. The RTL
description of the circuitry described in the last section (with an
hardware description language) has been written
and synthesized at gate-level using \emph{Synopsys Design Compiler} with 
a generic technology library (GTECH).
Considering the specific layout of each single logic element (NAND,
full-adder, latch etc.), it has been possible to get a rough estimation
of the number of transistors necessary to implement \disr{} logic.
Figure~\ref{fig:synthesys} shows the results of synthesis in terms of
number of devices (CNFETs) versus the number of network nodes while
the network scales up from $10\times10$ to $100\times100$ nodes.

@@

In particular, Figure~\ref{fig:synthesys} shows that the proposed 
implementation occupies about 17,5\% of the node budget. 
The main contribution is due to the controller circuitry followed by LED registers. 
Further, more than the
absolute number of devices itself, it is interesting to observe that
the complexity of the circuitry necessary to implement \disr{} increases
with a slowly growing trend. This is due to the
relatively simple logic of \disr{} which is almost coded with
scalable storage structures. For example, the number of registers
implementing the \emph{link\_visited[]} and \emph{link\_tvisited[]}
table follow the logarithmic function $N_{reg}=N_{port} \cdot
log_2(N)$ where \emph{Nport} is the number of the router’s ports and N is
the number of network’s nodes.

In order to have an idea of the maximum working clock frequency for
the implemented circuitry, timing results can be obtained from a 
gate-level netlist. Since a complete technology library useful for
commercial synthesis tool (e.g. Design Compiler) is not available for
this particular technology, we have
synthesized the RTL description of \disr{}, with a standard 32~nm CMOS
library, obtaining a delay of about $\mathrm{\tau_{\disr{}}=10} \ FO4$
(fan-out of four). Considering the results obtained
in~\cite{deng_isscc07}, which reports the ratio in terms of FO4
between a standard 32~nm CMOS technology and a carbon nanotube one, 
a rough delay estimation can be calculated. In particular in such work 
has been reported a ratio of:  
\begin{equation}
\mathrm{\frac{FO4_{CMOS}}{FO4_{CNFET}}=2}  
\end{equation}
Considering that F04 CMOS is about 30~ps, and the results of last
equation, a FO4 for the CNFET case is equal to about 15~ps, the working 
frequency can be computed as:
\begin{equation}
\mathrm{f_{clk}=\frac{1}{T_{max}}=\frac{1}{\tau_{\disr{}} \cdot FO4_{CNFET}}}=
\frac{1}{10\cdot 15 \cdot 10^{-12}}=6,6 GHz
\end{equation}
where $\mathrm{\tau_{\disr{}}}$ is expressed in terms of FO4.

With regard to the power consumption of the whole set of \disr{} devices: 
due to the fact that \disr{} circuitry will be active 
only once during system startup, an accurate power analysis related to switching 
activity is not provided in this work. After the setup phase, 
when all segments are mapped, the \disr{} block will stop its activity 
passing all the information related to segments to the routing
algorithm. For this reason the dynamic power falls to zero and
the only consumption is due to leakage effects. A comparison between 
the proposed circuitry and the state of the art implementation for the RPF 
algorithm, presented in~\cite{patwardhan_nanoarch06}, can show that there 
is not an appreciable discrepancy in terms of transistors cost and power 
consumption. In facts, RPF requires an overhead of about 1692 CNFET for a 
$100 \times 100$ network. Furthermore, regarding the power consumption, 
also the RPF approach has a setup phase beyond which, the consumption tends 
to zero.

### 2014-09-09 PAPER/introduction.tex
Fixed bio for each author and some other typos

\section{Introduction and Motivation}

Exploring long term alternatives to the CMOS technology is gaining
more and more relevance as the scaling trend of such devices keeps
introducing new challenges: power density, defect tolerance, testing
costs and wire delays are only a few of the many critical aspects
involved~\cite{itrs13}. 
While software parallelism and multicore
approaches~\cite{horowitz2004} and \cite{powell2009} are partially mitigating the impact of such
constraints on performances, it is likely that the growing computing demand will
eventually need even more radical architectural modifications and new
paradigms in order to address the Computer Design challenges of the next
decades.

In the last years, self-assembled nanoscale architectures~\cite{yan2003}
emerged as a promising technology due their tera/peta scale of
integration, defect tolerance and huge potential computing
capabilities. These technologies are certainly still at
their early stage of development, however different laboratory demos and
prof-of-concepts architectures have been presented~\cite{patwardhan2004, patwardhan2006_1, pistol2009}.
%, patwardhan2006
The main idea behind this approach is exploiting the physical regularity and
stability of DNA structures in order to create a scaffold onto which
nano-devices (e.g. nanowires and CNFETs~\cite{bachtold2001, tans1998, cui2001}) can be
attached. This can be achieved by designing appropriate complementary DNA tags for
each terminal to be placed, so that a nano device will be attached
only where its own DNA tag matches a complementary tag on the DNA grid
scaffold (see Figure~\ref{fig:nana}-a).
A detailed description of the chemical properties involved is far
beyond the scope of this paper (see also~\cite{braun1998, seeman1999}), so we will focus on
the three main challenges that this new fabrication process introduces
in Computer Design: \emph{(i) limited node complexity}, \emph{(ii) large scale
randomness} and \emph{(iii) high defect rates}.

The \emph{limited node complexity} aspect is directly
related to the use of complementary DNA tags in order to place circuit
components. In a traditional CMOS process the complexity is introduced
using a photolithography mask, so that larger (more complex) circuits
only require larger masks; conversely, self-assembly achieves
complexity by increasing the number of unique DNA tags, since more
different tags means having more control on component placement. Ideally, by specifying a single
and unique tag for each nano transistor terminal, we could exactly
choose where each component would be placed. But the number of DNA
symbols forming the DNA sequence is limited (sequence of 4 nucleobases
G,A,T,C) and so creating many different tags (of a given length) would
mean make them more similar to each other, increasing the probability
of incorrect/partial matching. To avoid this problem we should limit
the number of unique tags, thus limiting the complexity at each node.
In particular, a budget of about $10,000$ CNFETs per node has been
estimated in~\cite{liu_jetcs}.

\emph{Large scale randomness} is the other fundamental condition of
self-assembled technology: DNA tags allow a control placement inside
node grid, but there is no control of where each of these  grids will
be placed on the whole network. As a consequence, other typical
properties of regular networks cannot be guaranteed, e.g. being
connected to a fixed number of neighbors, having a determined
orientation and so on.

Finally, \emph{defect rates}: while they are very hardly tolerated in mask
based top down design, the same nature of a bottom up self-assembly
 process cannot assume such a deterministic device placement
process, thus defect tolerance is something like a design requirement
more than an exception to be avoided.

\begin{figure*}
\centering
\begin{tabular}{cc}
    \includegraphics[width=0.40\textwidth]{pictures/dna2b.eps} &
    \includegraphics[width=0.40\textwidth]{pictures/dna1_complex2.eps} \\
 (a) & (b)
 \end{tabular}
  \caption{(a) DNA tag matching and (b) resulting irregular topology}
  \label{fig:nana}
\end{figure*}
These aspects of a DNA self-assembly process lead to some important
implications to be addressed when approaching to the design of
a computer architecture: computational model should be based on a
distributed network of small processing and storage nodes, randomly
placed and interconnected (see Figure~\ref{fig:nana}-b).
Proof-of-concept of such architectures can be found for
example in~\cite{patwardhan2006_1}, where instruction and data
operands are moved around the network in order to be processed.
Independently of the particular policy chosen to route packets,
such networks must be able to guarantee deadlock freedom without any
prior assumption on topology.

In this work, we introduce \disr{}, a Distributed Segment-based approach
to deadlock freedom in large scale DNA self-assembled networks. Our
contribution aims to achieve the classical properties of a
segment-based approach~\cite{mejia_ipdps06} without requiring any
topology graph, external defect map or centralized algorithm
execution.  So the \disr{} approach is not intended to discover the
``optimal'' segment choice (ideally reachable with the knowledge of
the topology graph) but just to demonstrate a concrete model that can
fit into such complex, irregular and large sized networks.
It is important to
underline that, although we can assume an architecture similar to the
one depicted in Figure~\ref{fig:nana} as background, the topology discovery
problem faced in this work is more general and not strictly dependent on any of the
underlying computing architectures.

The paper is organized as follows: in Section~\ref{sec:related_works}
we summarize the main approaches for topology agnostic routing. Next, in
Section~\ref{sec:disr_concepts} are described the main elements and
the execution model of the proposed approach. In Section~\ref{sec:simulation},
simulations are carried out with the open-source tool Nanoxim to
demonstrate how \disr{} preserves some of the main segment-based
properties and how compares to the tree-based approaches. Finally a
draft of an architectural implementation is shown in
Section~\ref{sec:implementation} to evaluate the feasibility of the
approach in terms of node complexity.

\section{Background and Contribution}
\label{sec:related_works}
Several works addressed the problem of topology agnostic deadlock
freedom introducing virtual channels~\cite{sancho2002, skeie2002,
skeie2004, koibuchi2003}, but while they show interesting
performances, the low resources available to prevent deadlocks makes them not suitable for
the DNA self-assembled networks we are assuming as
scenario.
Other topology independent algorithms~\cite{schroeder1991,
koibuchi2001, cherkasova1996} do not require any virtual channel but
exploit the concept of ``turn prohibition''.
In particular, authors of~\cite{Patwardhan05evaluatingthe} exploit the creation of a spanning tree of the
topology, placing bidirectional restrictions to avoid that a packet
can traverse a given link in both up and down directions.
The hierarchical nature of this approach can lead to uneven traffic
distribution, with many packets traversing upper links (near to the
root), but this is quite acceptable in classical wide area networks
topologies with a limited number of nodes. Other approaches such as
FX~\cite{sancho2000} mitigate this
issue, but the set of turn restrictions is still prefixed
strictly depending on the particular tree root selected. 
Other solutions try to approach the issue of irregular
topologies by limiting the number~\cite{duato1997, gomez2004, koibuchi2008} or the
location~\cite{zhang2008, flich2008, liu2011} of missing links. This
restriction is clearly unacceptable given the hight-defect rates of
DNA self-assembled networks scenario. For the same reason, we also
avoid considering solutions based on hardware-redundancy to
dynamically recover defects as in~\cite{constantinides2006, kohler2010,  park2006, ebrahimi2013}.

@@

The \disr{} approach presented in this work is a first attempt to
achieve the same goals of a segment based turn prohibition without
requiring neither the knowledge of a topology graph nor centralized execution. In particular
\disr{} implements segment discovery and turn prohibition using a
completely different mechanism, based on a distributed exchange of
small setup packets requiring on average $15\%$ of the available
resource budget. The following main features distinguish \disr{}
from classic approaches to segment partitioning: 
\begin{itemize}
\item No centralized entity is globally responsible and/or aware of
what is going on, i.e. the status of the \disr{} execution is
collectively distributed among the nodes.
\item  No defect map and/or topology graph is used as input,
thus, the topology has to be discovered \emph{while} segments are
created.
\item At the end of the execution, no segment list is created or
stored anywhere: each node is only aware of belonging to a segment,
ignoring the presence of other nodes in the same segment and also the
presence of other segments in the network.
\end{itemize}

>>>>

\section{Introduction and Motivation}
\label{sec:introduction}

Exploring long term alternatives to the CMOS technology is gaining
more and more relevance as the scaling trend of such devices keeps
introducing new challenges: power density, defect tolerance, testing
costs and wire delays are only a few of the many critical aspects
involved~\cite{itrs13}. 
While software parallelism and multicore
approaches~\cite{horowitz2004,powell2009} are partially mitigating the impact of such
constraints on performances, it is likely that a growing computing demand will
eventually need even more radical architectural modifications and 
paradigms to address the Computer Design challenges of the next
decades.

In the last years, self-assembled nanoscale
architectures~\cite{yan2003} emerged as a promising technology due
their tera/peta scale of integration, defect tolerance and huge
potential computing capabilities. These technologies are certainly
still at their early stage of development, however different
laboratory demos and proof-of-concepts have been
presented~\cite{patwardhan2004, patwardhan2006, patwardhan2006_1,
pistol2009}. For a complete survey on self-assembled architectures and
how they compare against classical many-core systems in terms of
 technologies, fault tolerance, performance and software tools see
 also~\cite{Future}.
%, patwardhan2006
The main idea behind this approach is exploiting the physical
regularity and stability of DNA structures in order to create a
scaffold onto which nano-devices (e.g. nanowires and
CNFETs~\cite{bachtold2001, tans1998, cui2001}) can be attached. This
can be achieved by designing appropriate complementary DNA tags for
each terminal to be placed, so that a nano device will be attached
only where its own DNA tag matches a complementary tag on the DNA grid
scaffold (see Figure~\ref{fig:nana}-a).  Of course, a detailed
description of the chemical properties involved is far beyond the
scope of this paper (see also~\cite{braun1998, seeman1999}).

In this work, we present \disr{} (Distributed Segment Routing), a distributed
approach focused on addressing the three main challenges that this new fabrication process introduces
in nanoscale Network-on-Chip Design: \emph{(i) limited node complexity}, \emph{(ii) large scale
randomness} and \emph{(iii) high defect rates}.  
The \emph{limited node complexity} aspect is directly
related to the use of complementary DNA tags in order to place circuit
components. The traditional CMOS process introduces complexity 
with larger photolithography masks: more complex (larger) circuits
will only require larger masks. Conversely, a self-assembly process achieves
complexity by increasing the number of unique DNA tags, since having more
different tags means having more control on component placement. Ideally, by specifying a single
and unique tag for each nano device terminal, we could exactly
choose where each component would be placed in the design, but the number of DNA
symbols forming the DNA sequence is limited (sequence of 4 nucleobases
G,A,T,C) and so creating many different tags (of a prefixed length) would
mean make them more similar to each other, increasing the probability
of incorrect/partial matching. To avoid this problem 
the number of unique tags must be limited, this limits also the
complexity available at each node.
In this work a budget of about $10,000$ CNFETs per node
has been assumed, as estimated in~\cite{liu_jetcs}.
\emph{Large scale randomness} is the other fundamental condition of
self-assembled technology: DNA tags allow control placement inside the
node grid, but there is no control over the placement of these grids
in the whole network. As a consequence, other typical
properties of regular networks cannot be guaranteed, e.g. being
connected to a fixed number of neighbors, having a determined
orientation and so on.
Finally, \emph{defect rates}: while they are very hardly tolerated in mask
based top-down design, the same nature of a bottom-up self-assembly
 process cannot assume such a deterministic device placement
process, thus defect tolerance is more a design requirement
than an exception to be avoided.

\begin{figure*}
\centering
\begin{tabular}{cc}
    \includegraphics[width=0.45\textwidth]{pictures/dna2b.eps} &
    \includegraphics[width=0.45\textwidth]{pictures/dna1_complex2.eps} \\
 (a) & (b)
 \end{tabular}
  \caption{(a) Strands of complementary DNA tags self assemble to
  generate larger structures (b) The resulting irregular topology of
  processing(P) and storage(M) elements }
  \label{fig:nana}
\end{figure*}
These aspects of a DNA self-assembly process lead to some important
implications to be addressed when approaching to the design of
a nanoscale Network-on-Chip architecture: computational model should be based on a
distributed network of small processing and storage nodes, randomly
placed and interconnected (see Figure~\ref{fig:nana}-b).
Proof-of-concept of such architectures can be found for
example in~\cite{patwardhan2006_1}, where instruction and data
operands are moved around the network in order to be processed.
In addition, independently of the particular policy chosen to route instruction and
data packets, such networks must be able to guarantee deadlock freedom without any
prior assumption about the topology.

In this work, we introduce \disr{}, a Distributed Segment-based
approach to topology discovery and deadlock freedom in large scale DNA
self-assembled on chip networks. Our contribution aims to achieve the
classical properties of a segment-based approach~\cite{mejia_ipdps06}
without requiring any topology graph, external defect map or
centralized algorithm execution.  The \disr{} approach is not
intended to discover the ``optimal'' segment choice (ideally reachable
with the knowledge of the topology graph) but just to demonstrate a
concrete model that can fit into such complex, irregular and large
sized networks. It is important to underline that, although we can
assume as background an architecture similar to the one depicted in
Figure~\ref{fig:nana}, the \disr{} topology discovery
approach presented this work is more general and not strictly
dependent on any underlying computational model adopted.

The paper is organized as follows: in Section~\ref{sec:related_works}
we summarize the main approaches for topology agnostic deadlock
freedom. Next, in
Section~\ref{sec:disr_concepts} are described the main elements and
the execution model of the proposed approach. Further details about node behavior
in relation to the execution model are provided in Section~\ref{sec:execution_model}. 
In Section~\ref{sec:simulation},
simulations are carried out with the open-source tool Nanoxim to
demonstrate how \disr{} preserves some of the main segment-based
properties and how compares to the tree-based approaches. Finally a
draft of an architectural implementation is shown in
Section~\ref{sec:implementation} to evaluate the feasibility of the
approach in terms of node complexity.

\section{Background and Contribution}
\label{sec:related_works}
Several works addressed the problem of topology agnostic deadlock
freedom introducing virtual channels~\cite{sancho2002, skeie2002,
skeie2004, koibuchi2003}, but while they show interesting
performances, the low resources available to prevent deadlocks makes them not suitable for
the DNA self-assembled networks we are assuming as
scenario.
Other topology independent algorithms~\cite{schroeder1991,
koibuchi2001, cherkasova1996} do not require any virtual channel but
exploit the concept of ``turn prohibition''.
In particular, authors of~\cite{Patwardhan05evaluatingthe} exploit the creation of a spanning tree of the
topology, placing bidirectional restrictions to avoid that a packet
can traverse a given link in both up and down directions.
The hierarchical nature of this approach can lead to an uneven traffic
distribution, with many packets traversing upper links (near to the
root), but this is quite acceptable in classical wide area networks
topologies with a limited number of nodes. Other approaches such as
FX~\cite{sancho2000} mitigate this
issue, but the set of turn restrictions is still prefixed
strictly depending on the particular tree root selected. 
Other solutions try to approach the issue of irregular
topologies by limiting the number~\cite{duato1997, gomez2004, koibuchi2008} or the
location~\cite{zhang2008, flich2008, liu2011} of missing links. This
restriction is clearly unacceptable given the hight-defect rates of
DNA self-assembled networks scenario. For the same reason, we also
avoid considering solutions based on hardware-redundancy to
dynamically recover defects as in~\cite{constantinides2006, kohler2010,  park2006, ebrahimi2013}.

@@

The \disr{} approach presented in this work is a first attempt to
achieve the same goals of a segment based turn prohibition
requiring neither the knowledge of a topology graph nor centralized execution. In particular,
\disr{} implements segment discovery and turn prohibition using a
completely different mechanism, based on a distributed exchange of
small setup packets requiring a limited and scalable portion of the available
resource budget. The following main features distinguish \disr{}
from classic approaches to segment partitioning: 
\begin{itemize}
\item No centralized entity is globally responsible and/or aware of
what is going on, i.e. the status of the \disr{} execution is
collectively distributed among the nodes.
\item  No defect map and/or topology graph is used as input,
thus, the topology has to be discovered \emph{while} segments are
created.
\item At the end of the execution, no segment list is created or
stored anywhere: each node is only aware of being part of a segment,
ignoring the presence of other nodes in the same segment and also the
presence of other segments in the network.
\end{itemize}

### 2014-09-09 PAPER/paper.tex
fixes
Added soul package to hilight text with \hl command
general fixing
English fixes
Removed some typos
Fixed bio for each author and some other typos

% Elsevier specific styles
%------------------------------------------------------------------------------
\documentclass[review]{elsarticle}
%\documentclass[3p]{elsarticle}
\usepackage{lineno,hyperref}

@@

%\bibliographystyle{elsarticle-num}
\section*{References}
\bibliography{bibliography}
%------------------------------------------------------------------------------

>>>>

% Elsevier specific styles
%------------------------------------------------------------------------------
\documentclass[preprint]{elsarticle}
\usepackage{lineno,hyperref}

@@

%\bibliographystyle{elsarticle-num}
\section*{References}
\bibliography{bibliography}
%------------------------------------------------------------------------------
\vspace{10 mm}

Vincenzo Catania received the Laurea degree in Electrical Engineering
from the University of Catania, Italy, in 1982. Until 1984, he was
responsible for testing microprocessor system at STMicroelectronics,
Catania. He is a full professor of computer
science. His research interests include performance and reliability
assessment in parallel and distributed system, VLSI design, low power
design, and fuzzy logic.

Andrea Mineo received the BSc and MSc degrees in Electronic
Engineering from the University of Catania, Italy, in 2010 and 2013,
respectively, and is currently pursuing the PhD degree in Systems,
Energy, Computer and Telecommunications Engineering at the University
of Catania. His current research interests are VLSI systems,
Network-on-Chip architectures and emerging interconnect technologies
for on-chip networks.

Salvatore Monteleone is a post-doctoral research assistant at the
Department of Electrical, Electronic and Computer Engineering -
University of Catania, Italy. He obtained the BSc (2007) and MSc
(2010) in Computer Engineering at University of Catania where he also
completed the PhD course in Communications and Computer Engineering
(2014). His interests comprehend cooperative systems, user-centric
systems and Network-on-Chip architectures.

Davide Patti received the Laurea and PhD degrees in computer
engineering from the University of Catania, Italy, in 2003 and 2007,
respectively. His research focuses on Design Space exploration of VLIW
systems, Network-on-Chip architectures, DNA self-assembled networks
and Human-computer interfaces. He is currently a research assistant
at University of Catania.
